\section{Backward Angle Neutron Detector}

The Backward Angle Neutron Detector (BAND) is placed at the top of the CVT cart upstream of
the CLAS12 target. It consists of 116 scintillator bars, arranged in 18 rows and 5 layers.
Four bars are missing in the bottom of the detector due to obstruction. The bars have a cross
section of $7.2 \times 7.2\,\mathrm{cm}^{2}$ and they are 164 and $202\,\mathrm{cm}$ long in the
upper region of BAND. In the bottom region the bars are divided into two shorter bars
$51\,\mathrm{cm}$ to have a hole for the beamline and target installation. All bars are read-out
on both ends by PMTs (Hamamatsu R7724 and ET9214) giving a total of 232 active channels. 

In front of the first active layer of BAND, a veto layer is installed with 24 bars read-out only
on one side. Therefore, the total number of channels for BAND is 256.
The PMTs are placed in the fridge field region of the solenoid, and due to this they are encased in
a cylindrical shielding made up by a 2-mm-thick layer of mu-metal.

In order to operate the PMTs, high voltages (typically in the range of 1500 V) are provided by a
multi-channel CAEN SYS4527 mainframe with 11 A15350 cards (24 channel each).
The signal of each PMT is sent to an 50/50 splitter.
From the splitter one signal is sent to flash-ADCs (250 VXS, 16 channels/board), while the other
signal is sent to discriminators used by HPS (16 channels/board).
The discriminated time signal then goes to a TDC (CAEN VX1190A, 128 channels/board, 100 ps/channel
resolution). The read-out system is installed in the racks on the south side of the Space Frame on
the target level. 
In total, the system consists of 16 flash-ADCs in one VXS crate, 16 discriminators and a TDC in a
VME crate, and 16 splitters. Furthermore, a signal distribution card for the flash-ADCs and trigger
interface boards is installed in the crates.

The laser calibration system consists of a Photonics STV-01E-140 picosecond pulse laser with a
wavelength of $355\,\mathrm{nm}$, several splitters, reference photodiode, and a fiber distribution
system. All of these components are in a sealed, light-tight box. The output of the laser is about
$1\,\mu\mathrm{J}$ per pulse at $0.3\,$ns width (FWHM) that is attenuated and distributed to all
fiber outputs that each have an output of about $200\,\mathrm{pJ}$. The fibers are connected via a
patch panel to each scintillator bar.

\subsection{Hazards} 

\indent
\subsubsection{Electrical Hazard}

The electrical hazard to personnel can come from the high voltage that powers the PMTs, which need
about 1500 V to function. 

\subsubsection{Fall Hazard}

Fall hazard from maintenance and testing operations of BAND that require the use of ladders to access
system elements in up to 1.5\,m. 

\subsubsection{Magnetic Field Hazard}

BAND is placed in the fringe field of the solenoid (50 - 100 gauss). A hazard may arise during maintenance
operations in case metallic tools are used and for people with cardiac pacemakers, other electrical medical
devices, or metallic implants.

\subsubsection{Laser Hazard}

The laser hazard comes from the laser calibration system and its connection with fibers to the scintillator
bars when work is done on BAND. The system itself is closed, light-tight and the output on each fiber is
$\approx 200\,\mathrm{pJ}$. This intensity is comparable to that of a LED, however, the $355\,\mathrm{nm}$
wavelength could be damaging to the human eye since it is invisible to the eye and the natural eye reflex
will not be triggered. 

\subsection{Mitigations}

Any work on this system must be covered by ePAS Permit to Work(s) (PTW).

\indent
\subsubsection{Electrical Hazard Mitigations} 

The electrical hazard associated with the HV system is mitigated by the use of properly rated RG-59 cables
that are terminated at the voltage dividers and the HV supplies.
The HV supplies are grounded to their electronics racks as well. 
The maximum current provided by the HV distribution boards is quite low ($<1\,\mathrm{mA}$). The HV system
is designed to shutdown any channels that show an over-current condition. The HV boards must not be accessed
during operation; during maintenance work, performed by trained personnel, the HV is turned off, and the
power supply is switched off by the power switch on the back of the crate.

\subsubsection{Fall Hazard Mitigations} 

Fall hazard mitigation consists of approriate training like fall protection training. For individuals using
ladders, they are required to take the appropriate ladder training.

\subsubsection{Magnetic Field Hazard Mitigations}

The magnetic field hazard must be mitigated for both personnel and detector components. Normally no servicing
work is to be done with the BAND detector when the solenoid is energized.
This mitigates any hazard associated with personnel working in a strong magnetic field environment. 
After all maintenance work is done on BAND, the area must be inspected and all ferromagnetic tools must be
removed before the field of the solenoid is ramped up again. This mitigates the hazard for the detector. 

\subsubsection{Laser Hazard Mitigations}

The laser calibration system is a closed system with fibers connected to the scintillator bars via a patch
panel. All of these connections are light-tight. Furthermore, no fiber must be disconnected from the system
while it is operating. While the laser intensity in each fiber is very small ($\approx 200\,\mathrm{pJ}$),
and comparable to that of a LED, the $355\,\mathrm{nm}$ wavelength could be damaging to the human eye.
Therefore, one should not look directly at the fiber output - eye protection must be worn. 

Warning labels are applied on the enclosure of the laser system, the patch panel, and detector frame. The
box containing the laser system is interlocked such that if the box is open the power is off. If the system
is not in use by trained personnel, it will be powered off. During maintenance work on the laser system,
performed by trained personnel, the laser system is turned off, disconnected from the power supply, and
locked and tagged. The procedures for maintenance work on the laser system can be found in the LOP.

\subsection{Responsible Personnel}
\indent

Individuals responsible for the BAND system are (see Table~\ref{tb:band}):

\begin{table}[!htb]
 \centering
 \begin{tabular}{|l|c|c|c|c|}
\hline
 Name           & Dept.  & Phone        & email                    & Comments \\ \hline
Expert on call  &        & 757-310-7198 &                          & 1st contact \\ \hline
F. Hauenstein   & Hall B & 757-746-3395 & \href{mailto:hauenst@jlab.org}{\nolinkurl{hauenst@jlab.org}}     & 2nd Contact \\ \hline
O. Hen          & MIT    &              & \href{mailto:hen@mit.edu}{\nolinkurl{hen@mit.edu}}               & 3nd Contact \\ \hline
L. Weinstein    & ODU    &              & \href{mailto:weinstein@jlab.org}{\nolinkurl{weinstein@jlab.org}} & 4th contact  \\ \hline
 \end{tabular}
\caption{Personnel responsible for the CLAS12 BAND Detector.} 
\label{tb:band}
\end{table}
