\documentclass[11pt]{report}
%\documentclass[11pt]{article}
\usepackage{tabularx,graphicx,layout}
\usepackage{hyperref}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage{enumitem}
\usepackage{pifont}
\usepackage{version}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}

\begin{document}
\title{ Experiment Safety Assessment Document (ESAD) \\ 
for the Hall B CLAS12 Commissioning Run}
\date{\today\\v1.0}
%
% LaTeX Version 12 GeV ESAD 
%
%
\maketitle
\tableofcontents

\chapter{Introduction}

This ESAD document describes the identified hazards of the experiment and the measures 
taken to eliminate, control, or mitigate them. This document is part of the CEBAF 
experiment review process as defined in
\href{http://www.jlab.org/ehs/ehsmanual/manual/3120.html}{Chapter 3120} of the Jefferson 
Lab EHS\&Q manual, and will start by describing general types of hazards that might be 
present in any of the JLab experimental halls.  The document then addresses the hazards 
associated with subsystems of the equipment in Experimental Hall~B and their mitigation.  
Responsible personnel for each item are also noted. In case of life threatening emergencies 
call 911 and then notify the guard house at x5822 so that the guards can help the 
responders.  This document does not attempt to describe the function or operation of the 
various subsystems. Such information can be found in the individual subsystem Operations 
Manuals.

\chapter{General Hazards}

\section{Radiation}
	
CEBAF's high intensity and high energy electron beam is a potentially lethal direct 
radiation source. It can also create radioactive materials that are hazardous even after 
the beam has been turned off. There are many redundant measures aimed at preventing 
accidental exposure to personnel by the beam or exposure to beam-associated radiation 
sources that are in place at JLab. The training and mitigation procedures are handled 
through the JLab Radiation Control Department (RadCon). RadCon can be contacted as follows: 
For routine support and surveys, or for emergencies after-hours, call the RadCon cell phone 
at 876-1743. For escalation of effort or for emergencies, the RadCon manager (Vashek Vylet) 
can be reached as follows: Office: 269-7551, Cell: 218-2733 or Home: 772-6098.

Radiation damage to materials and electronics is mainly determined by the neutron 
dose (photon dose typically causes parity errors and is easier to shield against). 
Commercial-off-the-shelf (COTS) electronics is typically robust up to neutron doses of 
about $10^{13}$~n/cm$^2$. If the experimental equipment dose as calculated in the RSAD 
is beyond this damage threshold, the experiment needs to add an appendix on ``Evaluation 
of potential radiation damage" in the experiment-specific ESAD. There, the radiation 
damage dose, potential impact to equipment located in areas above this damage threshold, 
as well as the mitigating measures to be taken should be described.

\section{Fire}

Each of the experimental halls contains numerous combustible materials and flammable gases. 
In addition, they contain potential ignition sources, such as electrical wiring and equipment. 
General fire hazards and procedures for dealing with these are covered by JLab emergency 
management procedures. The JLab fire protection manager (Ed Douberly) can be contacted at 
269-6638.

\section{Electrical Systems}

Hazards associated with electrical systems are the most common risk in the experimental 
halls. Almost every subsystem requires AC and/or DC power. Due to the high current and/or 
high voltage requirements of many of these subsystems, they and their power supplies are 
potentially lethal electrical sources. In the case of superconducting magnets the stored 
energy is so large that an uncontrolled electrical discharge can be lethal for a period of 
time even after the actual power source has been turned off. Anyone working on electrical 
power in the experimental halls must comply with 
\href{http://www.jlab.org/ehs/ehsmanual/manual/6200.html}{Chapter 6200} of the Jefferson Lab 
EHS\&Q manual and must obtain approval of one of the responsible personnel. The JLab 
electrical safety point-of-contact (Todd Kujawa) can be reached at 269-7006.

\section{Mechanical Systems}

There exist a variety of mechanical hazards in all experimental halls at JLab. Numerous 
electro-mechanical subsystems are massive enough to produce potential fall and/or crush 
hazards.  In addition, heavy objects are routinely moved around within the experimental 
halls during reconfigurations for specific experiments. 

Use of ladders and scaffolding must comply with 
\href{http://www.jlab.org/ehs/ehsmanual/manual/6132.html}{Chapter 6132} of the Jefferson 
Lab EHS\&Q manual. Use of cranes, hoists, lifts, etc. must comply with
\href{http://www.jlab.org/ehs/ehsmanual/manual/6141.html}{Chapter 6141} of the Jefferson
Lab EHS\&Q manual. Use of personal protective equipment to mitigate mechanical hazards, 
such as hard hats, safety harnesses, and safety shoes, are mandatory when deemed necessary.
The JLab technical point-of-contact (Suresh Chandra) can be contacted at 269-7248.

\section{Strong Magnetic Fields}

Powerful magnets exist in all JLab experimental halls. Metal objects may be attracted 
by the magnet fringe field and become airborne, possibly injuring body parts or striking 
fragile components, resulting in a cascading hazard condition. Cardiac pacemakers or other 
electronic medical devices may no longer function properly in the presence of magnetic 
fields. Metallic medical implants (non-electronic) may be adversely affected by magnetic 
fields. Loss of information from magnetic data storage devices such as tapes, disks, and 
credit cards may also occur. Contact Jennifer Williams at 269-7882, in case of questions 
or concerns.

\section{Cryogenic Fluids and Oxygen Deficiency \\Hazard}

Cryogenic fluids and gases are commonly used in the experimental halls at JLab. When 
released in an uncontrolled manner these can result in explosion, fire, cryogenic burns, 
and the displacement of air, resulting in an oxygen deficiency hazard (ODH) condition. 
The hazard level and associated mitigation are dependent on the subsystem and cryogenic 
fluid. However, they are mostly associated with cryogenic superconducting magnets and 
cryogenic target systems. Flammable cryogenic gases used in the experimental halls 
include hydrogen and deuterium, which are colorless, odorless gases and, hence, not easily 
detected by human senses. Hydrogen air mixtures are flammable over a large range of relative 
concentrations from 4\% to 75\% H$_2$ by volume. Non-flammable cryogenic gases typically 
used include He and nitrogen. Contact Henry Robertson at 269-7285 or Mathew Wright at 
269-7722 in case of questions or concerns.

\section{Vacuum and Pressure Vessels}

Vacuum and/or pressure vessels are commonly used in the experimental halls. Many 
of these have thin aluminum or Kevlar/Mylar windows that are close to the entrance 
and/or exit of the vessels or beampipes. These windows burst if punctured accidentally 
or can fail if significant over-pressure were to exist. Injury is possible if a failure 
were to occur near an individual. All work on vacuum windows in the experimental halls 
must occur under the supervision of appropriately trained JLab personnel. Specifically, 
the scattering chamber and beamline exit windows must always be leak checked before 
service. Contact Will Oren 269-7344 for vacuum and pressure vessel issues.

\section{Hazardous Materials}

Hazardous materials in the form of solids, liquids, and gases that may harm people or 
property exist in the JLab experimental halls. The most common of these materials 
include lead, beryllium compounds, and various toxic and corrosive chemicals. Material 
Safety Data Sheets (MSDS) for hazardous materials in use in the experimental hall are 
available from the hall Safety Wardens. These are being replaced by the new standard
Safety Data Sheets (SDS) as they become available in compliance with the new OSHA 
standards. Handling of these materials must follow the guidelines of the EHS\&Q manual. 
Machining of lead or beryllium, which are highly toxic in powdered form, requires prior 
approval of the EHS\&Q staff. Lead Worker training is required in order to handle lead in 
the experiment halls. In case of questions or concerns, the JLab hazardous materials 
specialist (Jennifer Williams) can be contacted at 269-7882.

\section{Lasers}

High power lasers are often used in the experimental areas for various purposes. Improperly 
used lasers are potentially dangerous. Exposure to laser beams at sufficient power levels 
may cause thermal and photochemical injury to the eye, resulting in retina burns and 
blindness. Skin exposure to laser beams may induce pigmentation, accelerated aging, or 
severe skin burns. Laser beams may also ignite combustible materials creating a fire hazard. 
At JLab, lasers with power higher than 5~mW (Class IIIB) can only be operated in a controlled 
environment with proper eye protection and engineering controls designed and approved for the 
specific laser system. Each specific laser system shall be operated under the supervision of 
a Laser System Supervisor (LSS) following the Laser Operating Safety Procedure (LOSP) for 
that system approved by the Laboratory Laser Safety Officer (LSO). The LSO (Bert Manzlak) can 
be reached at 269-7556.
%
% details for each of the Halls
%
\chapter{Hall~B-Specific Equipment}

\section{Overview}

The Hall~B subsystems included in this chapter are considered part of the End-station 
Experimental Equipment for the CLAS12 Commissioning Run. Many of these subsystems impose 
similar hazards, such as those induced by magnets and magnet power supplies, high voltage 
systems, cryogenic systems, and vacuum systems. Note that a specific subsystem may have 
many unique hazards associated with it. For each major system, the hazards, mitigations, 
and responsible personnel are noted.

The material in this chapter is a subset of the material in the Hall~B subsystem Operations 
Manuals and is only intended to familiarize people with the hazards and responsible personnel 
for these systems. In no way should it be taken as sufficient information to use or operate 
this equipment.

\section{Detector Checkout System}

The Hall~B equipment readiness for the beam run will be done using the CEBAF Hot Checkout 
system (HCO)~\cite{hco}. Each detector system has its own subsystems and items that have 
to be checked and signed off by the appropriate groups. The systems that have been checked 
and signed as ready for beam will be indicated by a green ``thumbs up". Subsystems that are 
still waiting for readiness check-up and approval will be indicated with red crosses. 

In order to make sure that hall equipment that should be tied into the machine Fast Shutdown 
(FSD) system has been properly checked, the hall Work Coordinator must be notified by e-mail 
prior to the end of each installation period by the system owner that the checks have been 
performed in conjunction with accelerator operations (i.e. checking that the equipment's 
signals will in fact cause an FSD). These notifications will be noted in the Work Coordinator's 
final checklist as having been done. System owners are responsible for notifying the Work 
Coordinator that their system has an FSD tie-in so it can be added to the checklist.

The hall will get permission to run beam only after all systems tied to the delivery of the 
beam to the designated destination are signed off. At that point the main Hall~B system will 
indicate ``Ready" in the HCO tree.

\input{beamline_vacuum_systems}

\input{target_system}

\input{gas}

\input{daq}

\input{htcc}

\input{dc}

\input{ltcc}

\input{ftof}

\input{electromagnetic_calorimeter}

\input{ctof}

\input{silicon_tracker}

\input{cnd}

\input{mm}

\input{rich}

\input{ft}

\input{solenoid}

\input{torus}

\begin{thebibliography}{99}
\bibitem{hco} CEBAF Hot Check-Out db page (accessible from inside JLab only)
URL:https://accweb.acc.jlab.org/hco/

\bibitem{epics} 
EPICS Documentation. URL: http://www.epics.org/. 
See also http://www.aps.anl.gov/asd/controls/epics/EpicsDocumention\\/WWWPages/EpicsDoc.html

\bibitem{ehs} 
JLab EHS\&Q Manual. URL: http://www.jlab.org/ehs/ehsmanual/.

\bibitem{pss} 
JLab/CEBAF Personnel Safety System (PSS) Manual. URL: http://www.jlab.org/eng/ssg/user\_info.html

\bibitem{ops} 
Accelerator Operations Directive. \\
URL: http://www.jlab.org/ops\_docs/online\_document\_files/ACC\_online\_files/\\accel\_ops\_directive.pdf (URL is available inside JLab site). 

\end{thebibliography}

\end{document}

