\section{Low Threshold Cherenkov Counter}

The CLAS12 Low Threshold Cherenkov Counter (LTCC) is composed of four identical detectors
installed in the S2, S3, S5, and S6 positions upstream of FTOF on the Forward Carriage. 
The detectors are filled with C$_4$F$_{10}$ gas supplied by the Hall-B Gas system. The gas 
is cleaned, re-circulated, and maintained at a pressure between $1 - 4$~inches of wc with 
gas flow controllers and bubble pressure relief units. Each sector contains 36 PMTs energized 
by a HV power supply. Each PMT produces two outputs, connected to VME electronics (FADCs, 
TDCs) on the Forward Carriage.

\subsection{Hazards} 

There are three hazards identified with operation of the LTCC system. 
\begin{itemize}
\item Electrical hazard when the HVPS is energized for the PMTs.
\item Fall hazards from using man-lifts or ladders to access system elements during 
maintenance and testing operations. 
\item Gas pressure hazards when the detector is pressurized with C$_4$F$_{10}$, typically 
$1 -4$~inches of wc.
\end{itemize}

\subsection{Mitigations}

The HV hazard is mitigated by the maximum current settings on the power supply.

Harness training, man-lift training, ladder training, and fall protection training provides 
mitigation for the fall hazard during the detector maintenance.

Detector pressure and vacuum is limited to a maximum of $4$~inches of wc by the bubbler 
pressure relief units.

\subsection{Responsible Personnel}

Individuals responsible for the LTCC system are:

\begin{table}[!htb]
\centering
\begin{tabular}{|c|c|c|c|c|} \hline
Name           & Dept.& Phone        & email & Comments \\ \hline
Expert on call &      & 757-344-7174 &       & 1st contact \\ \hline
M. Ungaro      & JLab & 757-269-7578 &\href{mailto:ungaro@jlab.org}{\nolinkurl{ungaro@jlab.org}}&2nd contact \\ \hline
 \end{tabular}
\caption{Personnel responsible for the CLAS12 LTCC system.} 
\label{tb:ltcc}
\end{table}

