\subsection{PRad vaccuum chamber}
PRad experiment a large 5m long vacuum chamber extending from the target to the PRad detector system. 
There is a 1.7m diameter 63 mil Al window at one end of the vacuum chamber, just before the PRad detectors. 

\subsubsection{Hazards}
When this chamber is under vacuum, it has a very large stored energy. 
The accidental rupture of the window releases this stored energy, which presents a hazard to the personnel and equipment.

\subsubsection{Mitigation}

A window cover has been fabricated from 1/8” thick aluminum to protect the window from damage caused by something falling into it. This cover will be attached to the window at all times except when the experiment is running.  
\begin{enumerate}
\item Place barriers at all entries to the level 1 spaceframe.
\item Place signs at all entries to level 1 that warn personnel of a thin window under vacuum.
\item Place signs at all entries to level 1 that state earplugs and safety glasses must be worn to enter level 1.
\item Place earplugs and safety glasses at the entries to level 1.
\item Personnel authorized by the Hall Work Coordinator will pump down the vacuum tank and check for leaks.
\item The signs and barriers will remain in place if the tank is under vacuum.
\end{enumerate}

The protective window cover will be installed or removed only when no vacuum is in the tank.  This will remove the stored energy in the tank so people can work near the window.  
Any operations near the window, including maintenance and/or repair work on the PRAD detector, must be done only when there is no vacuum in the tank and protective covers are in place.
In the event of a vacuum system leak, personnel authorized by the Hall Work Coordinator will bleed up the vacuum tank vacuum.
If there is an emergency and personnel can enter the area, the vacuum tank can be bled up by personnel authorized by the Hall Work Coordinator.  If personnel cannot enter the area, the signs are in place to warn emergency responders of the thin window under vacuum
